---
title: "Introduction to metricsgraphics"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

[metricsgraphics](http://github.com/hrbrmstr/metricsgraphics) is an [htmlwidget](https://github.com/ramnathv/htmlwidgets) interface to the [MetricsGraphics.js](http://metricsgraphicsjs.org) JavaScript/D3 chart library.

The current `htmlwidget` wrapper for it does not provide support for MetricsGraphics histograms but does have basic support for it's best feature - time series charts.

### Installation

`devtools::install_github("hrbrmstr/metricsgraphics")`

### Usage

Building `metricsgraphics` charts follows the "piping" idiom made popular through the [magrittr](https://github.com/smbache/magrittr), [ggvis](http://ggvis.rstudio.com/) and [dplyr](http://github.com/dplyr) packages. This makes it possible to avoid one giant function with a ton of parameters and facilitates breaking out the chart building into logical steps. While MetricsGraphics.js charts may not have the flexibility of `ggplot2`, you can build functional, interactive [multi-]line, scatterplot & bar charts (histograms coming soon) and even link charts together.

All plots begin with `mjs_plot`, which sets up the widget. You then use `mjs_line`, `mjs_bar` or `mjs_point` to specify the "geom" you want to use for plotting. However, unlike `ggplot2` (or even `base` `plot`) you cannot combine "geoms". The only exception to that is adding more lines to a `mjs_line` plot. This is not a limitation of the package, but more a design principle of the MetricsGraphics JavaScript library.

### Examples

#### Basic Line Chart

This example shows a basic line chart with MetricsGraphics.js `baseline` [[1](https://github.com/mozilla/metrics-graphics/wiki/Graphic#wiki-baselines)] & `marker` [[2](https://github.com/mozilla/metrics-graphics/wiki/Graphic#wiki-markers)] annotations:

```{r}
library(htmltools)
library(htmlwidgets)
library(metricsgraphics)
library(RColorBrewer)

tmp <- data.frame(year=seq(1790, 1970, 10), uspop=as.numeric(uspop))

tmp %>%
  mjs_plot(x=year, y=uspop) %>%
  mjs_line() %>%
  mjs_add_marker(1850, "Something Wonderful") %>%
  mjs_add_baseline(150, "Something Awful")
```

#### Basic Bar Chart

```{r}
tmp %>%
  mjs_plot(x=uspop, y=year, width=500, height=400) %>%
  mjs_bar() %>%
  mjs_axis_x(xax_format = 'plain')
```

```{r}
```

#### Scatterplots

```{r}

mtcars %>%
  mjs_plot(x=wt, y=mpg, width=600, height=500) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")
```

```{r}

mtcars %>%
  mjs_plot(x=wt, y=mpg, width=600, height=500) %>%
  mjs_point(color_accessor=cyl,
            x_rug=TRUE, y_rug=TRUE,
            size_accessor=carb,
            size_range=c(5, 10),
            color_type="category",
            color_range=brewer.pal(n=11, name="RdBu")[c(1, 5, 11)]) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon") %>%
  mjs_add_legend(legend="X")
```

```{r}
mtcars %>%
  mjs_plot(x=wt, y=mpg, width=600, height=500) %>%
  mjs_point(least_squares=TRUE) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")
```

#### Muti-line charts

```{r}

set.seed(1492)
stocks <- data.frame(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4))

stocks %>%
  mjs_plot(x=time, y=X) %>%
  mjs_line() %>%
  mjs_add_line(Y) %>%
  mjs_add_line(Z) %>%
  mjs_axis_x(xax_format="date") %>%
  mjs_add_legend(legend=c("X", "Y", "Z"))
```

#### Linked Charts

```{r}
stocks2 <- data.frame(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4))

s1 <- stocks %>%
  mjs_plot(x=time, y=X, linked=TRUE, width=350, height=275) %>%
  mjs_line() %>%
  mjs_add_line(Y) %>%
  mjs_add_line(Z) %>%
  mjs_axis_x(xax_format="date") %>%
  mjs_add_legend(legend=c("X", "Y", "Z"))

s2 <- stocks2 %>%
  mjs_plot(x=time, y=X, linked=TRUE, width=350, height=275) %>%
  mjs_line() %>%
  mjs_add_line(Y) %>%
  mjs_add_line(Z) %>%
  mjs_axis_x(xax_format="date") %>%
  mjs_add_legend(legend=c("X", "Y", "Z"))
```

```{r eval=FALSE}
tagList(HTML("<table width='100%'><tr><td>"), 
        s1, 
        HTML("</td><td>"), 
        s2, 
        HTML("</td></tr></table>"))
```

```{r results='asis', echo=FALSE}
tagList(HTML("<table width='100%'><tr><td>"), s1, HTML("</td><td>"), s2, HTML("</td></tr></table>"))
```

```{r results='asis', echo=FALSE}
tagList(div(style="border-top:1px solid black; padding-top:10px; padding-bottom:20px; font-size:11px", 
            HTML("metricsgraphics R package by Bob Rudis<br/>"),  
            HTML("htmlwidgets R package by Ramnath Vaidyanathan, Kenton Russell<br/>"),
            HTML("MetricsGraphics JavaScript library by Ali Almossawi, Hamilton Ulmer and the Metrics Team at Mozilla."))
)
```
