---
title: "README"
author: "Bob Rudis"
date: January 04, 2015
output:
  md_document:
    variant: markdown_github
---

metricsgraphics is an 'htmlwidget' interface to the MetricsGraphics.js D3 chart library.

The current `htmlwidget` wrapper for it is minimaly functional and does not provide support
for metricsgraphics histograms and also does not really take advantage of metricsgraphics'
best feature - time series charts. Both will be added in the next release and support for 
legends will appear shortly thereafter.

Charts look best in a Boostrap page (unless you customize your own CSS).

The following functions are implemented:

- `mjs_plot`:  Create a new metricsgraphics.js plot
- `mjs_line`:  metricsgraphics.js linechart "geom"
- `mjs_point`:	metricsgraphics.js scatterplot "geom"
- `mjs_bar`:  metricsgraphics.js bar chart "geom"
- `mjs_axis_x`:  Configure x axis ticks & limits
- `mjs_axis_y`:	Configure y axis ticks & limits
- `mjs_labs`:	Configure axis labels & plot description

### News

- Version 0.1 released

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/metricsgraphics")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r eval=FALSE}
library(metricsgraphics)

tmp <- data.frame(year=seq(1790, 1970, 10), uspop=as.numeric(uspop))

tmp %>%
  mjs_plot(x=year, y=uspop) %>%
  mjs_line() %>%
  mjs_labs(title="Populations Recorded by the US Census", 
           description="This data set gives the population of the United States (in millions) as recorded by the decennial census for the period 1790–1970.")

tmp %>%
  mjs_plot(x=year, y=uspop, width=600) %>%
  mjs_line(area=TRUE) %>%
  mjs_labs(title="Populations Recorded by the US Census", 
           description="This data set gives the population of the United States (in millions) as recorded by the decennial census for the period 1790–1970.")

tmp %>% 
  mjs_plot(x=year, y=uspop, width=300, height=400) %>%
  mjs_bar() %>%
  mjs_labs(title="Sample Bar Chart")

mtcars %>% 
  mjs_plot(x=wt, y=mpg, width=400, height=300) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon", title="Scatterplot w/color & size")


mtcars %>% 
  mjs_plot(x=wt, y=mpg, width=400, height=300) %>%
  mjs_point(least_squares=TRUE) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon", title="Scatterplot w/least squares")

```


