---
title: "README"
author: "Bob Rudis"
date: January 05, 2015
output:
  md_document:
    variant: markdown_github
---

metricsgraphics is an 'htmlwidget' interface to the MetricsGraphics.js D3 chart library.

The current `htmlwidget` wrapper for it is minimaly functional and does not provide support
for metricsgraphics histograms and also does not really take advantage of metricsgraphics'
best feature - time series charts. Both will be added in the next release and support for 
legends will appear shortly thereafter.

Charts look best in a Boostrap page (unless you customize your own CSS).

You can see what the output below produces [on RPubs](http://rpubs.com/hrbrmstr/metricsgraphics0-1).

The following functions are implemented:

- `mjs_plot`:  Create a new metricsgraphics.js plot
- `mjs_line`:  metricsgraphics.js linechart "geom"
- `mjs_point`:	metricsgraphics.js scatterplot "geom"
- `mjs_bar`:  metricsgraphics.js bar chart "geom"
- `mjs_axis_x`:  Configure x axis ticks & limits
- `mjs_axis_y`:	Configure y axis ticks & limits
- `mjs_labs`:	Configure axis labels & plot description
- `mjs_add_baseline`: Sets a baseline line/label
- `mjs_add_marker`: Sets a marker line/label

### News

- Version 0.1 released
- Version 0.2 released - added support for markers & baselines + minimal support for time-series

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/metricsgraphics")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r eval=FALSE}
library(metricsgraphics)

tmp <- data.frame(year=seq(1790, 1970, 10), uspop=as.numeric(uspop))

tmp %>%
  mjs_plot(x=year, y=uspop) %>%
  mjs_line() %>%
  mjs_add_marker(1850, "Something Wonderful") %>%
  mjs_add_baseline(150, "Something Awful")


tmp %>%
  mjs_plot(x=year, y=uspop, width=600) %>%
  mjs_line(area=TRUE)

tmp %>% 
  mjs_plot(x=uspop, y=year, width=500, height=400) %>%
  mjs_bar() %>%
  mjs_axis_x(xax_format = 'plain')

mtcars %>% 
  mjs_plot(x=wt, y=mpg, width=400, height=300) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")


mtcars %>% 
  mjs_plot(x=wt, y=mpg, width=400, height=300) %>%
  mjs_point(least_squares=TRUE) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")


set.seed(1492)
dat <- data.frame(date=seq(as.Date("2014-01-01"), 
                           as.Date("2014-01-31"), 
                           by="1 day"),
                  value=rnorm(n=31, mean=0, sd=2))

dat %>%
  mjs_plot(x=date, y=value) %>%
  mjs_line() %>%
  mjs_axis_x(xax_format = "date")

```

