---
title: "Introduction to metricsgraphics"
author: "Bob Rudis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to metricsgraphics}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
[MetricsGraphics.js](http://metricsgraphicsjs.org/) is a library built on top of [D3](http://d3js.org/) that is optimized for visualizing and laying out time-series data. It provides a simple way to produce common types of graphics in a principled, consistent and responsive way. The library currently supports line charts, scatterplots and histograms as well as features like rug plots and basic linear regression.

The `metricsgraphics` R package wraps most of the functionality of MetricsGraphics.js into an [htmlwidget](http://htmlwidgets.org/). This package will replicate many of [the MetricsGraphics.js examples](http://metricsgraphicsjs.org/examples.htm) to demonstrate the functionality of the package.

```{r echo=FALSE}
suppressPackageStartupMessages(library(metricsgraphics))
suppressPackageStartupMessages(library(jsonlite))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(dplyr))
```

```{r}
library(metricsgraphics)
library(jsonlite)
library(RColorBrewer)
library(htmltools)
library(dplyr)

# this lets us add a title to the plot since the package follows the guidance
# of the htmlwidgets authors and does not include the MetricsGraphics.js title
# option to ensure consistent div sizing.

show_plot <- function(plot_object, title) {
  div(style="margin:auto;text-align:center", strong(title), br(), plot_object)
}
```

This is a simple line chart. You can remove the area portion by changing `area=TRUE` to `area=FALSE`. NOTE that you do _not_ have to use `show_plot` for normal `metricsgraphics` usage. It's used here for example purposes only.

```{r}
fake_users_1 <- fromJSON("http://metricsgraphicsjs.org/data/fake_users1.json")
fake_users_1$date <- as.Date(fake_users_1$date)

fake_users_1 %>%
  mjs_plot(x=date, y=value, width=600, height=200) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(area=TRUE) %>% 
  show_plot("Line Chart")

```

This is an example of a graphic with a confidence band and extended x-axis ticks enabled. NOTE that by specifying the `xaxformat="date"` we do not, then, have to pre-convert `date` with `as.Date()`

```{r}
confidence_band <- fromJSON("http://metricsgraphicsjs.org/data/confidence_band.json")

confidence_band %>%
  mjs_plot(x=date, y=value, format="percentage", width=600, height=200) %>%
  mjs_axis_x(xax_format="date", 
             show_secondary_x_label=FALSE, 
             extended_ticks=TRUE) %>% 
  mjs_line() %>% 
  mjs_add_confidence_band() %>% 
  show_plot("Confidence Band")
```

When we have a data object of integers and a small range of values, the auto-generated set of y-axis ticks are filtered so that we don't include fractional values.

```{r}
small_range <- fromJSON("http://metricsgraphicsjs.org/data/small-range.json")

small_range %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(interpolate="basic", area=TRUE) %>% 
  show_plot("Small Range of Integers")
  
```

The two graphics in this section are linked together. A rollover in one causes a rollover in the other. NOTE that setting `linked=TRUE` links _all_ the charts on
a given page/document that have that setting. This is a limitation of the MetricsGraphics.js library, not the package.

```{r}
brief_1 <- fromJSON("http://metricsgraphicsjs.org/data/brief-1.json")
brief_2 <- fromJSON("http://metricsgraphicsjs.org/data/brief-2.json")

brief_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200, linked=TRUE) %>%
  mjs_axis_x(xax_format="date", xax_count=4) %>% 
  mjs_line(area=TRUE) -> mjs_brief_1

brief_2 %>% 
  mjs_plot(x=date, y=value, width=600, height=200, linked=TRUE) %>%
  mjs_axis_x(xax_format="date", xax_count=4) %>% 
  mjs_line() -> mjs_brief_2

div(style="margin:auto;text-align:center", 
    strong("Linked Graphic"), br(), mjs_brief_1,
    strong("Other Linked Graphic"), br(), mjs_brief_2)
```

Handling a solitary data point. NOTE that date+time objects (e.g. `POSIXct`) are a TODO for `metricsgraphics`.


```{r}
solitary <- data.frame(
  date=as.Date("2015-03-05"),
  value=12000
)

solitary %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_point() %>% 
  show_plot("Singleton")
```

This line chart contains multiple lines. NOTE that the example build a data frame from the MetricsGraphics.js example JSON data. Unlike icky straight JavaScript constructs, we get to work with our own sane, native data types.

```{r}
fake_users2_list <- fromJSON("http://metricsgraphicsjs.org/data/fake_users2.json")
fake_users2 <- data.frame(
  date=fake_users2_list[[1]]$date,
  value_1=fake_users2_list[[1]]$value,
  value_2=fake_users2_list[[2]]$value,
  value_3=fake_users2_list[[3]]$value
)

fake_users2 %>% 
  mjs_plot(x=date, y=value_1, width=600, height=200) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_line() %>% 
  mjs_add_line(value_2) %>% 
  mjs_add_line(value_3) %>% 
  mjs_add_legend(c("Line 1", "Line 2", "Line 3")) %>% 
  show_plot("Multi-Line Chart")

fake_users2 %>% 
  mjs_plot(x=date, y=value_1, width=600, height=200) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(color="blue") %>% 
  mjs_add_line(value_2, color="rgb(255,100,43)") %>% 
  mjs_add_line(value_3, color="#ccccff") %>% 
  mjs_add_legend(c("Line 1", "Line 2", "Line 3")) %>% 
  show_plot("Multi-Line Char with Custom Colors")
```

Since we pass in whole data frames, there is no "missing" data points or columns. Also, since a linked chart example was shown already, we can't show that multi-line example, so all three will not be reproduced here.

We _can_ show inline legend labels (placed to the right) vs bottom.

```{r}
fake_users3_list <- fromJSON("http://metricsgraphicsjs.org/data/fake_users3.json")
fake_users3 <- data.frame(
  date=fake_users3_list[[1]]$date,
  value_1=fake_users3_list[[1]]$value,
  value_2=fake_users3_list[[2]]$value,
  value_3=fake_users3_list[[3]]$value
)

fake_users3 %>% 
  mjs_plot(x=date, y=value_1, width=600, height=200, right=40) %>%
  mjs_axis_x(xax_format="date") %>% 
  mjs_line() %>% 
  mjs_add_line(value_2) %>% 
  mjs_add_line(value_3) %>% 
  mjs_add_legend(c('US', 'CA', 'DE'), inline=TRUE) %>% 
  show_plot("Labeling Lines")
```

This is an example of a graphic where we're not plotting dates on the x-axis and where the axes include labels and the line animates on load. We've also enabled extended ticks along the y-axis.

```{r}
xnotondate <- fromJSON("http://metricsgraphicsjs.org/data/xnotdate.json")

xnotondate %>% 
  mjs_plot(x=males, y=females, width=600, height=240, 
           left=80, right=40, bottom=50) %>% 
  mjs_line(animate_on_load=TRUE, area=FALSE) %>% 
  mjs_labs("Males", "Females") %>% 
  mjs_axis_y(extended_ticks=TRUE) %>% 
  show_plot("Axis Labels")
```

Here is an example that shows percentages.

```{r}
some_percentages <- fromJSON("http://metricsgraphicsjs.org/data/some_percentage.json")

some_percentages[[1]] %>% 
  mjs_plot(x=date, y=value, format="percentage", width=600, height=200) %>% 
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(area=TRUE) %>% 
  show_plot("Some Percentages")
```

Here is an example that uses custom units for currency.

```{r}
some_currency <- fromJSON("http://metricsgraphicsjs.org/data/some_currency.json")

some_currency %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date") %>% 
  mjs_line() %>% 
  mjs_axis_y(yax_units="$") %>% 
  show_plot("Some Currency")
```

You can change the y-axis' scale to logarithmic by setting `y_scale_type` to `"log"`.

```{r}
log_scale <- fromJSON("http://metricsgraphicsjs.org/data/log.json")

log_scale %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(area=TRUE) %>% 
  mjs_axis_y(y_scale_type="log") %>% 
  show_plot("Log Scale")
```

You can hide either axis by setting `show=FALSE` in `mjs_axis_x` or `mjs_axis_y`.

```{r}
fake_users_1 <- fromJSON("http://metricsgraphicsjs.org/data/fake_users1.json")
brief_1 <- fromJSON("http://metricsgraphicsjs.org/data/brief-1.json")

fake_users_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date", show=FALSE) %>% 
  mjs_line() -> no_x
  
brief_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date") %>%
  mjs_axis_y(show=FALSE) %>% 
  mjs_line() -> no_y

div(style="margin:auto;text-align:center", 
    strong("No X Axis"), br(), no_x,
    strong("No Y Axis"), br(), no_y)
```

You can now change line color without resorting to CSS:


```{r}
fake_users_1 <- fromJSON("http://metricsgraphicsjs.org/data/fake_users1.json")

fake_users_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date", show=FALSE) %>% 
  mjs_line(color="#8c001a", area=TRUE) %>% 
  mjs_axis_y(rug=TRUE) %>% 
  show_plot("Colors!")
```

You can set rug plots on either axis by setting `rug=TRUE` in `mjs_axis_x` or `mjs_axis_y`.

```{r}
fake_users_1 <- fromJSON("http://metricsgraphicsjs.org/data/fake_users1.json")

fake_users_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date", show=FALSE) %>% 
  mjs_line() %>% 
  mjs_axis_y(rug=TRUE) %>% 
  show_plot("Rug Plots")
```

Markers are vertical lines that can be added at arbitrary points. Markers that are close to each other won't collide. NOTE that `metricsgraphics` will do date conversions from strings automagically.

```{r}
some_percentages <- fromJSON("http://metricsgraphicsjs.org/data/some_percentage.json")

some_percentages[[1]] %>% 
  mjs_plot(x=date, y=value, format="percentage", width=600, height=200) %>% 
  mjs_axis_x(xax_format="date") %>% 
  mjs_line(area=TRUE) %>% 
  mjs_add_marker("2014-02-01", "1st Milestone") %>% 
  mjs_add_marker(as.Date("2014-03-15"), "2nd Milestone") %>% 
  show_plot("Markers")
```

Baselines are horizontal lines that can added at arbitrary points.

```{r}
fake_users_1 <- fromJSON("http://metricsgraphicsjs.org/data/fake_users1.json")

fake_users_1 %>% 
  mjs_plot(x=date, y=value, width=600, height=200) %>% 
  mjs_axis_x(xax_format="date", show=FALSE) %>% 
  mjs_add_baseline(160000000, "a baseline") %>% 
  mjs_line(area=TRUE) %>% 
  show_plot("Baselines")
```

Scatterplots are first-class citizens as well

```{r}
points_1 <- fromJSON("http://metricsgraphicsjs.org/data/points1.json")

points_1 %>% 
  mjs_plot(x=x, y=y, width=600, height=460) %>% 
  mjs_point(y_rug=TRUE) %>% 
  mjs_axis_x() %>% 
  show_plot("Simple Scatterplot")
```

```{r}
points_1 %>% 
  mjs_plot(x=x, y=y, width=600, height=460) %>% 
  mjs_point(y_rug=TRUE, color_accessor=v, color_type="category", color_range=c("green", "orange")) %>% 
  mjs_axis_x() %>% 
  show_plot("Color mapping")
```

```{r}
points_1 %>% 
  mjs_plot(x=x, y=y, width=600, height=460) %>% 
  mjs_point(y_rug=TRUE, x_rug=TRUE, color_accessor=z, size_accessor=w, color_type="category") %>% 
  mjs_axis_x(rug=TRUE) %>% 
  show_plot("Size Too!")
```

This does not mimic the MetricsGraphics.js examples but shows off `mjs_grid` functionality:

```{r}
moar_plots <- lapply(1:7, function(x) {
  mjs_plot(rbeta(10000, x, x), width="250px", height="250px", linked=TRUE) %>%
    mjs_histogram(bar_margin=2) %>%
    mjs_labs(x_label=sprintf("Plot %d", x))
})

mjs_grid(moar_plots, nrow=4, ncol=3, widths=c(rep(0.33, 3)))
```
